\speculake: Hiding Malware in Speculative Execution

### 2018-08-06 paper/triggers.tex
pass on triggers

In addition, an adversary could use a benign application (like OpenSSL) to
\emph{communicate} information covertly to the malicious payload program. For
example, with OpenSSL, the attacker could have a pair of cipher uncommon cipher
suites, and using one results in communicating a 1-bit, while the other
communicates a 0-bit to the payload. To receive this, the payload would have to
do indirect jump patterns corresponding to both cipher suites, with the
corresponding speculative entry points shifting in the appropriate bit. Thus, an
adversary can communicate remotely (over a network) to the payload program
indirectly via a benign application.

We observe that communication could also go in the other direction: from
the payload program back to the remote adversary, also via a benign application
intermediary. The payload program could influence the performance of the benign
application, and the adversary could time responses from the benign application
to receive covert information from the malicious payload.

@@

In addition to using separate trigger programs, ExSpectre can also use
\emph{trigger inputs} to a payload program to initiate its malicious behavior.
While existing fuzzers and symbolic execution tools can discover traditional
input triggers, we can leverage other Spectre variants...

TODO: Jack

@@

TODO: Jack

>>>>

In addition, an adversary could use a benign application (like OpenSSL) to
\emph{communicate} information covertly to the malicious payload program. For
example, with OpenSSL, the attacker could have a pair of uncommon cipher
suites, where using one results in communicating a 1-bit, while use of the other
communicates a 0-bit to the payload. To receive data, the payload would have to
do indirect jump patterns corresponding to OpenSSL code for processing both cipher suites, with the
corresponding speculative entry points shifting in the appropriate bit. Thus, an
adversary can communicate remotely (over a network) to the payload program
indirectly via a benign application.

We observe that communication could also go in the other direction: from
the payload program back to the remote adversary, also via the benign application
intermediary. The payload program could influence the performance of the benign
application, and the adversary could time responses from the benign application
to receive covert information from the malicious payload~\cite{netspectre}. This would allow the
malicious payload to operate \emph{entirely speculatively}, without assistance
from the real world for keeping state.

@@

\FigSpectreOne

In addition to using separate trigger programs, ExSpectre can also use
\emph{trigger inputs} to a payload program to initiate its malicious behavior.
While existing fuzzers and symbolic execution tools can discover traditional
input triggers, we can leverage other Spectre variants to obfuscate our triggers.

To do this, we use Specualtive Buffer Overflows (SBO)~\cite{kiriansky2018speculative}
to redirect control flow
to a speculative entry point specified by user input.
We design a payload program that takes arbitrary user input and performs appropriate
bounds checks to ensure no traditional control flow violations could be exploited.
However, using the Spectre~1.1 variant, control flow can still be violated speculatively,
allowing the adversary to force a speculative entry point based entirely off the
input provided. This allows the trigger to be an input, potentially even
provided over a network if the payload program accepts network data. Traditional
symbolic execution and code coverage fuzzers will be unable to discover this trigger input, as
they do not model the speculative state of the CPU.

%such that evaluated code is safe, however a direct branch can
%be bypassed speculatively given significant lead up with consistent input 
%to train the branch predictor (see Figure~\ref{fig:spectre-one}). This allows the input
%to become the trigger.

To create an SBO-triggered payload program, we make the following code pattern,
as seen from Kiriansky and Waldspurger's Spectre~1.1
description~\cite{kiransky2018speculative}:
\begin{lstlisting}
    if (y < lenc)
        c[y] = z;
\end{lstlisting}
With user controlled 
\texttt{y} and \texttt{z} and sporadically uncached \texttt{lenc}, an attacker 
can \textit{speculatively} overflow array \texttt{c}
to overwrite a return address (or function pointer, etc) and redirect control
flow. Note that the bounds check on \texttt{y} will ensure that this program
will not actually allow a buffer overflow to occur, but the attacker can
nonetheless use this to influence a speculative entry point based on their
choice of \texttt{y} and \texttt{z}.

We make use of this pattern in a willing payload such that user input can intentionally
mistrain the branch predictor by repeatedly sending valid (in-bounds) values of
\texttt{y} before sending a value that would overflow the bounds of \texttt{c}.
The speculative entry point is also chosen by the 
trigger in this scenario as \texttt{z} contains the address of the 
speculative entry point, allowing the attacker to create a ROP-style speculative 
execution path through the payload. We describe our implementation of this in
Section~\ref{tk}.

@@

Speculative Store Bypass (SSB) (Spectre variant 4) can similarly be used to
construct an internal (input-based) trigger using the CPU's speculative
load-store forwarding~\cite{spec-store-bypass}. In a speculative store bypass,
the CPU incorrectly speculates that a store will not alias with a future load,
and uses a stale (wrong) value for the result of the speculative load.

To redirect control flow, a payload program could use a function pointer or
indirect branch target register as the destination for a speculative store
bypass, causing the CPU to use a stale value to speculatively determine where it
would go. The stale value could be controlled by a previous unrelated input,
allowing an adversary to specify the speculative entry point in a carefully
crafted data input. While the program never
executes at this stale address in reality, the CPU will briefly speculatively
execute there, enabling \speculake payloads. Like the speculative buffer
overflow, this trigger also allows ROP style chains to execute a series of
speculative gadgets.

%The speculative primitive is shared by all versions of Spectre as an upper
%bound on what can be done speculatively. What is required to adapt a spectre 
%vulnerability into the \speculake technique is an attacker controlled method of 
%triggering control flow redirection in a willing payload to a speculative entry point.
